---
title: "Intermediate Elective 2"
author: "Minh Tri Ngo"
format: pdf
---
# Set Up
## Packages
```{r}
#| label: packages

# general use and cleaning
library(tidyverse)
library(here)
library(janitor)
library(snakecase)

# wrapping axis labels
library(scales)

# reading in .xlsx files
library(readxl)

# visualizing missing data
library(naniar)

# arranging plots
library(patchwork)

# imported libraries from example
library(showtext)
library(camcorder)
library(ggtext)
library(nrBrand)
library(glue)
library(ggfx)

```

## Data
```{r}
# taxonomic information
taxon_list <- read_csv(here("data", "taxon_list.csv")) 

# aquatic invertebrates
aq_ins <- read_xlsx(here("data", "Aquatic Sampling Data-2026-03-10.xlsx"),
                    sheet = "Aquatic Insects") 

# site information
sites <- read_xlsx(here("data", "Aquatic Sampling Data-2026-03-10.xlsx"),
                   sheet = "sites")

# water quality information
water_quality <- read_xlsx(here("data", "Aquatic Sampling Data-2026-03-10.xlsx"),
                           sheet = "Water Quality",
                           na = c("", "n/a", "N/A", "over", "173+")) 


# creating new object from sites
ncos_sites <- sites |> 
  # filter to only include NCOS sites sampled four times a year
  filter(sector == "NCOS" & sampling_frequency == "Quarterly")



# creating new clean object from taxon_list
taxon_list_clean <- taxon_list |> 
  # converting taxon name to snake case (no spaces or capital letters,
  # only underscores)
  mutate(verbatim_name = to_any_case(verbatim_name, case = "snake"))

# creating new clean object from water quality
water_quality_clean <- water_quality |> 
  # clean column names
  clean_names() |> 
  # filtering join: only include sites in ncos_sites
  semi_join(ncos_sites, by = "site") |> 
  # create new column to join with later data frames
  unite("date_site", date, site, remove = TRUE) |> 
  # group by date_site column
  group_by(date_site) |> 
  # calculate median pH, salinity, DO
  summarize(med_ph = median(p_h, na.rm = TRUE),
            med_sal = median(salinity_ppt, na.rm = TRUE),
            med_do = median(dissolved_oxygen_mg_l, na.rm = TRUE)) |> 
  # ungroup data frame
  ungroup() |> 
  # filter out salinity outliers
  filter(date_site != "2022-08-12_NVBR") 


# creating new clean object from aquatic inverts
aq_ins_clean <- aq_ins |> 
  # clean column names
  clean_names() |> 
  # filtering join: only include sites occurring in ncos_sites
  semi_join(ncos_sites, by = c("site")) |> 
  # renaming columns
  rename(date = date_on_vial) |> 
  # select columns of interest
  select(date, sample_type, site,
         ostracod,
         copepod,
         hemiptera_corixidae_boatman,
         cladocera,
         nematode,
         diptera_ceratopogonidae, 
         annelida_oligochaete,
         diptera_chironomid,
         annelida_polychaete) |> 
  # filter to only include "filtered beaker" samples
  filter(sample_type %in% c("FB 250", "FB250")) |> 
  # convert the data frame to long format
  pivot_longer(cols = ostracod:annelida_polychaete,
               names_to = "taxon",
               values_to = "abundance") |> 
  # group by date, site, taxon
  group_by(date, site, taxon) |> 
  # calculate average abundance per liter (sum abundance divided by 7.5 liters)
  summarize(ave_lit = sum(abundance, na.rm = TRUE)/7.5) |> 
  # ungroup data frame
  ungroup() |> 
  # create a new column called `date_site` to join with water quality
  unite("date_site", date, site, remove = FALSE) |> 
  # join with water quality
  left_join(water_quality_clean, by = c("date_site")) |> 
  # join with taxon list
  left_join(taxon_list_clean, by = c("taxon" = "verbatim_name")) |> 
  # add full names of sites
  mutate(site_full = case_when(
    site == "NVBR" ~ "North of Venoco Bridge",
    site == "NEC" ~ "South of Venoco Bridge",
    site == "NMC" ~ "Main Channel",
    site == "NPB" ~ "South of Phelps Bridge",
    site == "NPB1" ~ "North of Phelps Bridge",
    site == "NPB2" ~ "Phelps Road",
    site == "NWP" ~  "West Pond",
    site == "NDC" ~ "Devereux Creek"
  )) |> 
  # setting factor levels for site
  mutate(site_full = fct_reorder(site_full, med_sal, .fun = "median", .na_rm = TRUE),
         site_full = fct_rev(site_full))

# creating new object from clean aquatic inverts 
copepods_df <- aq_ins_clean |> 
  # filter to only include copepods
  filter(taxon == "copepod") |> 
  # log + 1 transform mean abundance per liter
  mutate(log_ave_lit = log(ave_lit + 1)) 

# creating new clean object from aquatic inverts
ab_do <- aq_ins_clean |> 
  # log transform mean abundance per liter
  mutate(log_ave_lit = log(ave_lit+1),
         # convert taxon to title
         taxon = to_any_case(taxon, case = "title"),
         # order taxon levels by abudance
         taxon = fct_reorder(taxon, ave_lit, .fun="max"))
```

# 1. Explain Inspiration
- The visualization I chose for my inspiration was the "Can an exploding 
snowman predict the summer season" by Nicola Rennie.
- I felt that this visualization makes sense as I would like to analyze 
temperature in relation to the different taxon availability and the 
graph allows me to analyze and communicate my findings the msot efficiently.
- The variables that I will choose in my dataset to show my visualization 
will be dissolved, taxon, and average abundance. In terms of visual 
components, the dissolved will be on the x axis. Taxon will have different
colors to differentiate. And the y axis will have the average abundance.

# 2. Plan figure

# 3. Code figure
```{r}
#| label: import code

# Load fonts --------------------------------------------------------------

font_add("Poppins",
  regular = "fonts/Poppins/Poppins-Regular.ttf",
  italic = "fonts/Poppins/Poppins-Italic.ttf",
  bold = "fonts/Poppins/Poppins-Bold.ttf",
  bolditalic = "fonts/Poppins/Poppins-BoldItalic.ttf"
)
font_add("FA",
  regular = "fonts/Font Awesome 6 Free-Solid-900.otf"
)
showtext_auto()
showtext_opts(dpi = 300)
title_font <- "Poppins"
body_font <- "Poppins"


# Define colours and fonts-------------------------------------------------

bg_col <- "#F2F4F8"
text_col <- "#151C28"
highlight_col <- "cornflowerblue"


# Start recording ---------------------------------------------------------

gg_record(
  dir = file.path("2025", "2025-12-02", "recording"),
  device = "png",
  width = 7,
  height = 6,
  units = "in",
  dpi = 300
)


# Define text -------------------------------------------------------------

social <- nrBrand::social_caption(
  bg_colour = bg_col,
  icon_colour = highlight_col,
  font_colour = text_col,
  font_family = body_font
)
icon <- "<span style='font-size:8pt;font-family:\"FA\";color:#3C598E;'>&#xf111;</span> "
title <- "Can temperature predict average abundance?"
st <- glue("The Böögg is a snowman effigy made of cotton wool and stuffed with fireworks, created every year for Zurich's *Sechselaeuten* spring festival. The saying goes that the quicker the Böögg's head explodes, the finer the summer will be. The Böögg was {icon}<span style='color: #3C598E;'>**correct**</span>* only {perc}% of the time, so maybe just flip a coin?")
cap <- paste0(
  "**Note**: *Correct is defined as above or below median burning duration predicting below or above median temperatures.<br>**Data**: data.geo.admin.ch | sechselaeuten.ch<br>**Graphic**: ", social
)

longest_burn <- plot_data |>
  slice_max(duration)
hottest <- plot_data |>
  slice_max(tre200m0)

annot1 <- glue("**Longest burn time**<br>{longest_burn$duration} minutes in {longest_burn$year}")
annot2 <- glue("**Hottest summer**<br>{hottest$tre200m0}°C on average in {hottest$year}")

annot_data <- tribble(
  ~x, ~y, ~label,
  54, 16.7, annot1,
  16, 21.2, annot2
)

# Plot --------------------------------------------------------------------

ggplot(
  data = plot_data
) +
  annotate(
    "rect",
    xmin = median(plot_data$duration), xmax = Inf,
    ymin = median(plot_data$tre200m0), ymax = Inf, alpha = 0.5,
    fill = "grey"
  ) +
  annotate(
    "text",
    x = 60, y = 21.5, label = "Slow burn, hot summer",
    colour = "grey30", family = body_font, fontface = "bold.italic",
    hjust = 1.05
  ) +
  annotate(
    "rect",
    xmin = median(plot_data$duration), xmax = Inf,
    ymin = -Inf, ymax = median(plot_data$tre200m0), alpha = 0.5,
    fill = highlight_col
  ) +
  annotate(
    "text",
    x = 60, y = 15.5, label = "Slow burn, cool summer",
    colour = "#3C598E", family = body_font, fontface = "bold.italic",
    hjust = 1.05
  ) +
  annotate(
    "rect",
    xmin = -Inf, xmax = median(plot_data$duration),
    ymin = median(plot_data$tre200m0), ymax = Inf, alpha = 0.5,
    fill = highlight_col
  ) +
  annotate(
    "text",
    x = 0.5, y = 20.8, label = "Fast\nburn,\nhot\nsummer",
    colour = "#3C598E", family = body_font, fontface = "bold.italic",
    hjust = 0
  ) +
  annotate(
    "rect",
    xmin = -Inf, xmax = median(plot_data$duration),
    ymin = -Inf, ymax = median(plot_data$tre200m0), alpha = 0.5,
    fill = "grey"
  ) +
  annotate(
    "text",
    x = 0.5, y = 16.2, label = "Fast\nburn,\ncool\nsummer",
    colour = "grey30", family = body_font, fontface = "bold.italic",
    hjust = 0
  ) +
  with_outer_glow(
    geom_point(
      mapping = aes(
        x = duration, y = tre200m0,
        shape = correct, colour = correct
      ),
      size = 3
    ),
    colour = "white",
    expand = 3
  ) +
  scale_colour_manual(values = c("grey50", "#3C598E")) +
  geom_textbox(
    data = annot_data,
    mapping = aes(x = x, y = y, label = label),
    family = body_font,
    colour = text_col, fill = alpha(bg_col, 0.2),
    box.colour = "transparent"
  ) +
  scale_x_continuous(limits = c(0, 60)) +
  scale_y_continuous(limits = c(15, 22)) +
  labs(
    x = "Burning duration (minutes)",
    y = "Average air temperature 2m above ground (°C)",
    title = title, subtitle = st,
    caption = cap
  ) +
  coord_cartesian(expand = FALSE, clip = "off") +
  theme_minimal(base_size = 10, base_family = body_font) +
  theme(
    legend.position = "none",
    plot.margin = margin(10, 20, 10, 10),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    plot.background = element_rect(fill = bg_col, colour = bg_col),
    panel.background = element_rect(fill = bg_col, colour = bg_col),
    plot.title = element_textbox_simple(
      colour = text_col,
      hjust = 0,
      halign = 0,
      margin = margin(b = 0, t = 5),
      family = title_font,
      face = "bold",
      size = rel(1.5)
    ),
    plot.subtitle = element_textbox_simple(
      colour = text_col,
      hjust = 0,
      halign = 0,
      margin = margin(b = 35, t = 5),
      family = body_font
    ),
    plot.caption = element_textbox_simple(
      colour = text_col,
      hjust = 0,
      halign = 0,
      margin = margin(b = 0, t = 10),
      family = body_font
    ),
    strip.text = element_textbox_simple(
      face = "bold",
      margin = margin(t = 10),
      size = rel(0.9)
    ),
    panel.grid.minor = element_blank(),
    axis.title.x = element_text(hjust = 1, margin = margin(t = 5)),
    axis.title.y.left = element_text(
      angle = 0, vjust = 1.07,
      family = body_font,
      size = rel(0.9),
      margin = margin(r = -215)
    )
  )


# Save gif ----------------------------------------------------------------

ggsave(
  filename = file.path("2025", "2025-12-02", paste0("20251202", ".png")),
  width = 7,
  height = 6,
  bg = bg_col,
  units = "in",
  dpi = 300
)

gg_playback(
  name = file.path("2025", "2025-12-02", paste0("20251202", ".gif")),
  first_image_duration = 4,
  last_image_duration = 20,
  frame_duration = .25,
  background = bg_col
)
```


